---
title: "DM_partC"
author: "Group A-11"
output:
  word_document: default
  html_document: default
---

```{r setup, message=FALSE, eval = FALSE}
rm(list=ls())
library(rvest)
library(httr)
library(dplyr)
library(stringr)
library(tidyverse)
library(XML)
```


#The following code is intended do download all the required XML files using UK Food Standard Ageny's API. First, we assigned UK Food standard Agency's API url and imported it to R using read_html(). Then, we obtain links which contain the xml file of each city in the UK by specifying the specific html class.Create a vector of URLs for XML files for each city. Finally, We the XML files are saved into a dedicated folder.

```{r download XML, eval = FALSE, message=FALSE}


url_path_data<-"https://data.food.gov.uk/catalog/datasets/38dd8d6a-5ab1-4f50-b753-ab33288e3200"
url_path_source<-read_html(url_path_data)


url_path_source%>%
  html_nodes(".o-dataset-distribution--link" )

# take the link from 'href' in the class
url_path_source%>%
  html_nodes(".o-dataset-distribution--link" )%>%
  html_attr("href")


#extract the title
url_path_source%>%
  html_nodes(".c-dataset-element--title")%>%
  html_text()%>%
  trimws(.)

url_path_source%>%
  html_nodes(".o-dataset-distribution--link" )%>%
  html_attr("title")

urls_to_download<-url_path_source%>%
  html_nodes(".o-dataset-distribution--link" )%>%
  html_attr("href")

data_titles<-url_path_source%>%
  html_nodes(".c-dataset-element--title")%>%
  html_text()%>%
  trimws(.)


# we  create a dataframe to combine the vector
my_urls_download<-data.frame(
  title=data_titles,
  urls=urls_to_download
)

#change it into character
my_urls_download$titles<-as.character(my_urls_download$title)
my_urls_download$urls<-as.character(my_urls_download$urls)

my_urls_download<-my_urls_download[-1,] #remove the first row "!FHRS API Search Page"


for (url_index in 1:nrow(my_urls_download)) {
  url_here<-my_urls_download$urls[url_index]
  filename_vector<-strsplit(my_urls_download$urls[url_index],split="/")[[1]] #split the link and take the filename
  filename<-filename_vector[length(filename_vector)]
  download.file(url=url_here,
                destfile=paste0("partc_dataset/",filename))
  }

```

#The following code is meant to transform the XML files into a dataframe that captures the required information. First, an empty dataframe for the final structure called 'partc_df' was created. Next, the paths of the XML files were specified and imported as a list xmlToList() to convert it into a list. Then, XML leaves that contain the required information were specified, in this case, the 'EstablishmentCollection'. Afterwards, the values of the required fields were stored into a local dataframe which will only contain 1 row. Finally, by the end of every iteration the local dataframe is merged into the final dataframe, in other words, compounding the rows into the final dataframe, 'partc_df'.


```{r read XML files, eval = FALSE}

# setup empty df for the final solution
partc_df <- data.frame()

# get xml file paths 
folder_path <- 'partc_dataset'
file_path <-list.files(folder_path, pattern = "en-GB.xml")
xml_file_path <- paste(folder_path, file_path,sep="/")



for (k in 1:length(xml_file_path)){
  
  # Transform the XMLs into lists
  this_path <-xml_file_path[k]
  my_list <- XML::xmlToList(this_path)
  this_establishment_collections <- my_list$EstablishmentCollection

  for(i in 1:length(this_establishment_collections)){
  
    # harvest the values for a row
    this_df <- this_establishment_collections[i]$EstablishmentDetail
    
    frsid <- this_df$FHRSID
    LocalAuthorityBusinessID <- this_df$LocalAuthorityBusinessID
    BusinessName <- this_df$BusinessName
    BusinessType <- this_df$BusinessType
    BusinessTypeID <- this_df$BusinessTypeID
    AddressLine1 <- this_df$AddressLine1
    AddressLine2 <- this_df$AddressLine2
    AddressLine3 <- this_df$AddressLine3
    AddressLine4 <- this_df$AddressLine4
    PostCode <- this_df$PostCode
    RatingValue <- this_df$RatingValue
    RatingKey <- this_df$RatingKey
    RatingDate <- this_df$RatingDate
    LocalAuthorityCode <- this_df$LocalAuthorityCode
    LocalAuthorityName <- this_df$LocalAuthorityName
    LocalAuthorityWebSite <- this_df$LocalAuthorityWebSite
    LocalAuthorityEmailAddress <- this_df$LocalAuthorityEmailAddress
    Score_Hygiene <- this_df$Scores$Hygiene
    Score_Structural <- this_df$Scores$Structural
    Score_ConfidenceInManagement <- this_df$Scores$ConfidenceInManagement
    SchemeType <- this_df$SchemeType
    NewRatingPending <- this_df$NewRatingPending
    Geocode_Longitude <- this_df$Geocode$Longitude
    Geocode_Latitude <- this_df$Geocode$Latitude
  

    # create a local dataframe
    my_local_df <- tibble::tibble(
      frsid = ifelse(is.null(frsid) == TRUE, 'NULL', frsid),
      LocalAuthorityBusinessID = ifelse(is.null(LocalAuthorityBusinessID) == TRUE, 'NULL', LocalAuthorityBusinessID),
      BusinessName = ifelse(is.null(BusinessName) == TRUE, 'NULL', BusinessName),
      BusinessType = ifelse(is.null(BusinessType) == TRUE, 'NULL', BusinessType),
      BusinessTypeID = ifelse(is.null(BusinessTypeID) == TRUE, 'NULL', BusinessTypeID),
      AddressLine1 = ifelse(is.null(AddressLine1) == TRUE, 'NULL', AddressLine1),
      AddressLine2 = ifelse(is.null(AddressLine2) == TRUE, 'NULL', AddressLine2),
      AddressLine3 = ifelse(is.null(AddressLine3) == TRUE, 'NULL', AddressLine3),
      AddressLine4 = ifelse(is.null(AddressLine4) == TRUE, 'NULL', AddressLine4),
      PostCode = ifelse(is.null(PostCode) == TRUE, 'NULL', PostCode),
      RatingValue = ifelse(is.null(RatingValue) == TRUE, 'NULL', RatingValue), 
      RatingKey = ifelse(is.null(RatingKey) == TRUE, 'NULL', RatingKey),
      RatingDate = ifelse(is.null(RatingDate) == TRUE, 'NULL', RatingDate),
      LocalAuthorityCode = ifelse(is.null(LocalAuthorityCode) == TRUE, 'NULL', LocalAuthorityCode),
      LocalAuthorityName = ifelse(is.null(LocalAuthorityName) == TRUE, 'NULL', LocalAuthorityName),
      LocalAuthorityWebSite = ifelse(is.null(LocalAuthorityWebSite) == TRUE, 'NULL', LocalAuthorityWebSite),
      LocalAuthorityEmailAddress = ifelse(is.null(LocalAuthorityEmailAddress) == TRUE, 'NULL', LocalAuthorityEmailAddress),
      Score_Hygiene = ifelse(is.null(Score_Hygiene) == TRUE, 'NULL', Score_Hygiene),
      Score_Structural = ifelse(is.null(Score_Structural) == TRUE, 'NULL', Score_Structural),
      Score_ConfidenceInManagement = ifelse(is.null(Score_ConfidenceInManagement) == TRUE, 'NULL', Score_ConfidenceInManagement),
      SchemeType = ifelse(is.null(SchemeType) == TRUE, 'NULL', SchemeType),
      NewRatingPending = ifelse(is.null(NewRatingPending) == TRUE, 'NULL', NewRatingPending),
      Geocode_Longitude = ifelse(is.null(Geocode_Longitude) == TRUE, 'NULL', Geocode_Longitude),
      Geocode_Latitude = ifelse(is.null(Geocode_Latitude) == TRUE, 'NULL', Geocode_Latitude)
    )
  
  # replace 'NULL' value as NA
  my_local_df[my_local_df == 'NULL'] <- NA  
  
  # merge partc_df with the local dataframe
  partc_df <- dplyr::bind_rows(partc_df,my_local_df)
  }
  
}

print(nrow(partc_df))
```

