---
title: "data prepare"
author: "Yibo"
date: "2/18/2020"
output: html_document
---

```{r setup, include=FALSE}
knitr::opts_chunk$set(echo = TRUE)
```

```{r}
library(data.table)
library(readr)
library(tidytext)
library(tidyverse)
library(textcat)
library(RSQLite)
library(cld3)
```

```{r}
# For Amsterdam 
Amsterdam_listings <- fread("/Users/zhaoyibo/Desktop/TERM2/TA assignment/file/Amsterdam_listings.csv.gz")

Amsterdam_reviews <- fread("/Users/zhaoyibo/Desktop/TERM2/TA assignment/file/Amsterdam_reviews.csv.gz")
```

{r}
#create a database
airbnb_db <- dbConnect(SQLite(),dbname = "new_airbnb_db.db")
#store data in the sqlite
#table listing
dbWriteTable(airbnb_db, "listing", Amsterdam_listings,append = TRUE)
#table reviews
dbWriteTable(airbnb_db, "reviews", Amsterdam_reviews,append = TRUE)
a <- dbReadTable(airbnb_db,"listing")
dbListTables(airbnb_db)
b <- dbGetQuery(airbnb_db, "select  listing_id,id,date,reviewer_id,reviewer_name,comments from reviews")
#disconnect the database
dbDisconnect(airbnb_db)
names(Amsterdam_reviews)

```{r}
#Question1.a
#What are the dominant words per aggregation category (neighborhood, access to public transport etc.)?

#which variables we should analyze?(especially summary/space/description)
#Transit/Neighborhood’s/overview/Description/Access/Interaction/House rule/Host_about

reviews_variables <- c("reviewer_id","comments")
listings_variables <- c("host_about","house_rules","interaction","access","transit", "neighborhood_overview","description")

reviews_selected <- Amsterdam_reviews %>%
  select(reviews_variables)
reviews_selected_sample <- reviews_selected[1:1000,]

listings_selected <- Amsterdam_listings %>%
  select(listings_variables)
#detect the language
reviews_selected_sample$language <- detect_language(reviews_selected_sample$comments)
reviews_selected_sample <- filter(reviews_selected_sample,language == "en")  
reviews_corpus_source <- VectorSource(reviews_selected$comments) 
reviews_corpus <- VCorpus(reviews_corpus_source)
#Create the clean_corpus function
clean_corpus <- function(corpus){
  corpus <- tm_map(corpus, removePunctuation)
  corpus <- tm_map(corpus, content_transformer(tolower))
  corpus <- tm_map(corpus, removeWords, words = c(stopwords("en"),"amsterdam"))
  corpus <- tm_map(corpus, stripWhitespace)
  corpus <- tm_map(corpus, removeNumbers)
  return(corpus)
}
cleaned_reviews_corpus <- clean_corpus(reviews_corpus)
reviews_corpus_tdm <- TermDocumentMatrix(cleaned_reviews_corpus)
reviews_corpus_m <- as.matrix(reviews_corpus_tdm)

term_frequency <- rowSums(reviews_corpus_m)
term_frequency <- sort(term_frequency, decreasing = TRUE)
head(term_frequency,10)
#Question.b


```
